% \pdfbookmark[0]{Acknowledgement}{Acknowledgement}
\addchap{Acknowledgements}
\label{sec:acknowledgement}

A PhD, for which you need to do a research project, is a daunting task. How could I possibly frame the questions
that would lead to significant discoveries, design and interpret an experiment so that the conclusions are
absolutely convincing, foresee difficulties and find ways around them or, failing that, solve them when these
occurred?

These were the reasons why I did not apply for PhD projects after my masters in Mathematics but rather decided
to wander around the Himalayas. The seeds of this PhD were sown then, during my 3-year-long immersion with the
mountain communities of Ladakh. Living among mountains of sand, I recognized the futility of human life and was
also alarmed by how small human-scale actions could transform these mountains beyond recognition. How is this
even possible and what can be done about it? Looking back, those were the questions that pushed me to pursue
this journey. 

But why did I end up here in Switzerland, my first foreign country, and how did I get an opportunity to pursue a
PhD in the domain of glaciology, a subject which I had little knowledge about? Short answer: the trust and support
of three people: Sonam Wangchuk, Felix Keller, and Martin Hoelzle. With Sonam, great ideas were always 
around; all I had to do was to pick my favourite and pursue it. It was together with him that I began to interrogate
all things ice.  With Felix, Switzerland became a country where my second home is. Without his support, I cannot
imagine how I would have transitioned into the Swiss way of life. In Martin, I found a colleague twice my age
whose engagement around scientific questions was unmatched. I owe a lot to each of them who have mentored me
in Phyang, Samedan, and Fribourg for the past 7 years.

A PhD thesis does not write itself, nor does its author operate from a deserted island. As such, I would like to
acknowledge the many individuals that contributed to the realization of this work in, what will most likely be,
one of the most read sections of this thesis. This work would not have been possible without the support,
collaboration, and friendship of my family, friends, and colleagues. To that end, I attempt to thank all of them
in the paragraphs below.

First and foremost, words cannot express my gratitude and appreciation for professor Martin Hoelzle, my mentor.
You initiated and guided me through the wondrous world of science, at every step giving me that extra push to
believe in myself. Although not immediately clear to me at the beginning, being trained by you to be an academic
has been a privilege, and I am grateful for the entire journey: from writing proposals and manuscripts to
suffering through the destructive feedback together, from discussing research problems to helping others write
their bachelor's and master's theses, from designing experiments on the blackboard to establishing
Swiss and Indian ice stupa laboratories, from mere equations to real insights. To today, almost holding my PhD.

My PhD project was quite interdisciplinary, and for a while, whenever I ran into a problem, I pestered several people for help: from the univerity concierge Tony for his experience with pipelines to the faculty who
were experts in the various disciplines that I needed. I remember the day when Johannes Oerlemans (who won the
International Glaciological Society's Richardson medal 2 years later) told me he did not know how to solve the
problem I was facing in his area. I was a first-year graduate student and I figured that Oerlemans knew much
more than I did. If he did not have the answer, nobody did.

That is when it hit me: nobody did. That is why it was a research problem. And being my research problem, it was
up to me to solve it. Once I faced that fact, the going got easier. The crucial lesson was that the scope of things I
did not know was not merely vast but, for all practical purposes, infinite. That realization, instead of
discouraging, was liberating. If our ignorance is infinite, the only possible course of action is to muddle
through as best as we can.

Dozens of people have contributed with days and days of work for collecting the associated datasets. My special
thanks go to the Himalayan Institute of Alternatives, Ladakh (HIAL) team, who provided administrative and
material support to conduct measurement campaigns in India. Norboo Thinles, Nishant Tiku, Sourabh Maheshwari, and
Dr. Tom Matthews were instrumental for making these campaigns a success. Without the unconditional support of
Daniel Bürki from Guttannen Bewegt, the frequent drone flights and expensive installations during the past three
winters would not have been possible. I would also like to thank Adolf Kaeser and Mr. Flavio Catillaz from
Eispalast Schwarzsee. Even though no datasets from this site appear in the thesis or our publications, the field
experience we gained there was a necessary precondition for success in the sites that followed. 

In hindsight, my PhD alone would not have been able to reveal all the insights hidden in these datasets.
Therefore, I am grateful that Michelle Stirnimann and Maria Gruber opted to explore these datasets in their
master's and bachelor's theses work. Moreover, all the fieldwork remained lively thanks to their presence and
enthusiasm over this work.

One of the main reasons I arrived (almost) everyday to work are the fabulous colleagues of the Cryosphere
department, many of whom became good friends. My office mates over the years, Shafaq, Mario, Ottavia, Romain,
Rebecca, and Esther made our room quite lively, which I missed a lot during periods of remote work.
These colleagues coming from literally all parts of the world gave me an incredible international experience.
It was also a great experience to teach bachelor's courses with Horst and Eric. Luke deserves a special mention
as the most wise one in the department. His timely recommendation to use the Schwarzsee Eispalast as
my first experiment site kick-started my PhD in earnest. I would also like to thank the administrative staff for
welcoming me into the department. David and Sylvie, apart from taking care of my logistic and financial hurdles,
also nudged me in every encounter to become a better French speaker. Thank you Nicole, for taking that extra care towards
sorting all my visa anxieties. Alex showed me how cool technicians can be with his stories and youthfulness. I
feel very blessed to have spent time with all of you, ranging from many after-work drinks, numerous coffee
breaks, lunches, dinners, conferences, train rides, and treks. All of you contributed to the exceptional warm
and cosy atmosphere in our department, and I cherish every minute. 

Without the generous support from the Swiss Government Excellence Scholarship (SGES) and the University of
Fribourg, the scientific literature on ice stupas would likely have remained incomplete and I would have grown
old regretting that. Beyond financial support, the SGES were also very accomodating for my field work
requirements during a global pandemic and made exceptions for me that had never been made before. I am also
grateful to have received grants from the Swiss Polar Institute and GlaciersAlive Association that enabled the
extensive fieldwork over the past few winters. 

Not everything is science! Eluckkiya made sure it was that way. Without her, I feel I would have languished much
longer with the innumerable research questions emerging around this nascent topic. You not only provided me with
love and faith, but also showed me how to put goals in perspective to enjoy life beyond deadlines.
I can safely say that I would never have been able to finish my thesis in good sanity without your support, and
I am very grateful for that. 

Last but not least, I would like to express the most sincere gratitude to my parents. Some of my hard life
choices, like volunteering in the mountains or shifting abroad, were worth considering only because I always had
their financial and emotional support to lean back on.



